---
title: "Air BnB Data Analysis"
author: Shravani Samala
date: "4/2/2020 "
output: html_document
---

```{r, message=F, warning=F}
# front-matter
rm(list = ls()) #clear the workspace

library(ggplot2)
library(ggformula)
library(corrplot)
library(dplyr)
library(lmerTest)
library(lme4)
library(knitr)
library(tidyverse)
library(readr)
library(mosaic)
library(data.table)
library(VGAM)
library(lmtest)

```
# Introduction

*overall_satisfaction = rating on a 0-5 scale.
*satisfaction = 1 if overall_satisfaction is 5, 0 otherwise
*price = price for one night (in dollars)
*reviews = number of reviews posted
*room_type = Entire home/apt, Private room, or Shared room
*accommodates = number of people the unit can hold
*bedrooms = number of bedrooms
*minstay = minimum length of stay (in days)
*neighborhood = neighborhood where unit is located (1 of 43)
*district = district where unit is located (1 of 9)
*WalkScore = quality of the neighborhood for walking (0-100)
*TransitScore = quality of the neighborhood for public transit (0-100)
*BikeScore = quality of the neighborhood for biking (0-100)
*PctBlack = proportion of black residents in a neighborhood
*HighBlack = 1 if PctBlack above .60, 0 otherwise
```{r}
myFile2 <- "https://raw.githubusercontent.com/proback/BYSH/master/data/airbnb.csv"
AirBnb.data <- read_csv(myFile2)

head(AirBnb.data)
```


```{r}
AirBnb.data<-
AirBnb.data%>%
  rename(hotelID = X1)%>%
  mutate(priceHigh = factor(ifelse(price >= 100,1,0)),
         satisfaction = factor(ifelse(overall_satisfaction == 5.0, 1, 0))) # 1 means high price, 0 means low price
head(AirBnb.data)
glimpse(AirBnb.data)
View(AirBnb.data)
```

# Research Question 1: What variables contribute to a customer's complete satisfaction with a hotel? 
As seen in the summary statistics, the mean bedrooms, number fo people the unit accommodates, price, Walk score, Transit score, and Bike Score for being completely satisfied (satisfaction = 1) and not completely satsified (satisfaction = 0) are all very similar. The only one that looks significantly different is the price of the unit. After looking at the boxplots, the means for bedrooms, transit score, and bike score appear to be very similar and the means for number of people the unit accommodates, price, and walk score look different. Therefore, I will fit a logistic model to each of these variables to see if they are significant in predicting the odds of whether of not a customer is completely satisfied. 
```{r}
AirBnb.data %>%
  group_by(satisfaction) %>%
  summarise(mean(bedrooms), 
            mean(accommodates), 
            mean(price), 
            mean(minstay), 
            mean(WalkScore), 
            mean(TransitScore), 
            mean(BikeScore))

ggplot(AirBnb.data, aes(x = satisfaction, y = bedrooms)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = accommodates)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = price)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = minstay)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = WalkScore)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = TransitScore)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = BikeScore)) + geom_boxplot(color = "black") + coord_flip()


```


```{r}

# Bedtooms
bedModel <- glm(satisfaction ~ bedrooms, family = "binomial", data = AirBnb.data)
summary(bedModel)

# Accomodations
accModel <- glm(satisfaction ~ accommodates, family = "binomial", data = AirBnb.data)
summary(accModel)

# Price
priceModel <- glm(satisfaction ~ price, family = "binomial", data = AirBnb.data)
summary(priceModel)

# Minimum stay
minstayModel <- glm(satisfaction ~ minstay, family = "binomial", data = AirBnb.data)
summary(minstayModel)

# Walk Score
minstayModel <- glm(satisfaction ~ WalkScore, family = "binomial", data = AirBnb.data)
summary(minstayModel)

# Transit Score
minstayModel <- glm(satisfaction ~ TransitScore, family = "binomial", data = AirBnb.data)
summary(minstayModel)

# Bike Score
minstayModel <- glm(satisfaction ~ BikeScore, family = "binomial", data = AirBnb.data)
summary(minstayModel)

```

```{r}
#Full Model 
fullModel <- glm(satisfaction ~ (accommodates + price + WalkScore)^2, family = "binomial", data = AirBnb.data)

fullFinalModel <- step(fullModel, direction = "backward")
summary(fullFinalModel)
```

```{r}
redFinalModel <- glm(satisfaction ~ accommodates + price + WalkScore, family = "binomial", data = AirBnb.data)
summary(redFinalModel)

lrtest(fullFinalModel, redFinalModel)
```


# Research Question 2: To what extent do hotels within the same neighborhood have similar prices?

```{r}
modelA <- lmer(price ~ 1 + (1|neighborhood), data = AirBnb.data)
summary(modelA)

```