---
title: "Air BnB Data Analysis"
author: Shravani Samala
date: "4/2/2020 "
output: html_document
---

```{r, message=F, warning=F}
# front-matter
rm(list = ls()) #clear the workspace

library(ggplot2)
library(ggformula)
library(corrplot)
library(dplyr)
library(lmerTest)
library(lme4)
library(knitr)
library(tidyverse)
library(readr)
library(mosaic)
library(data.table)
library(VGAM)
library(lmtest)

```
# Introduction

My family and I have had to stay in many hotels over our lives. Personally, whenever I have gone on school trips, college visits, and many other types of travels, I would stay in a hotel with my friends and family. Sometimes, my mom would gather the whole family around her computer and ask us which hotel looked better to us. My dad worries about the price of the hotel, my mom reads all the reviews, and my sister and I just wonder if they serve breakfast in the morning. That being said, everyone's decision on which hotel to stay in depends on many factors. However, I have never stayed at an Air Bnb. Apparently, we can rent out entire homes or just a specific room of someone else's property. So, I am assuming the decisions people make about which Air Bnb to stay in is pretty similar to those they make when looking at hotels. 

First, I would like to know what variables most often lead to greater odds that a person is completely satisfied with their stay at an AirBnb (satisfaction = 1, overall_satisfaction = 5). Then, I would like to look at what factors contribute to the price of the Air Bnb. 

Variables included in data: 

*overall_satisfaction = rating on a 0-5 scale.
 
*price = price for one night (in dollars)

*reviews = number of reviews posted

*room_type = Entire home/apt, Private room, or Shared room

*accommodates = number of people the unit can hold

*bedrooms = number of bedrooms

*minstay = minimum length of stay (in days)

*neighborhood = neighborhood where unit is located (1 of 43)

*district = district where unit is located (1 of 9)

*WalkScore = quality of the neighborhood for walking (0-100)

*TransitScore = quality of the neighborhood for public transit (0-100)

*BikeScore = quality of the neighborhood for biking (0-100)

*PctBlack = proportion of black residents in a neighborhood

Variables created: 

*satisfaction = 1 if overall_satisfaction is 5, 0 otherwise

*priceHigh = 1 if price is greater than or equal to 120, 0 otherwise

*entire = 1 if room_type is Entire home/apt, 0 otherwise

*private = 1 if room_type is Private room, 0 otherwise

*shared = 1 if room_type is Shared room, 0 otherwise

*HighBlack = 1 if PctBlack above .60, 0 otherwise

*HighRev = 1 if unit has more than 100 reviews posted, 0 otherwise

### Importing and Cleaning the Data
```{r}
myFile2 <- "https://raw.githubusercontent.com/proback/BYSH/master/data/airbnb.csv"
AirBnb.data <- read_csv(myFile2)

head(AirBnb.data)
```


```{r}
AirBnb.data<-
AirBnb.data%>%
  rename(ID = X1)%>%
  mutate(priceHigh = factor(ifelse(price >= 120,1,0)), # 1 means high price, 0 means low price
        satisfaction = factor(ifelse(overall_satisfaction == 5.0, 1, 0)),
        entire = factor(ifelse(room_type == "Entire home/apt", 1, 0)), 
        private = factor(ifelse(room_type == "Private room", 1, 0)),
        shared = factor(ifelse(room_type == "Shared room", 1, 0)),
        HighBlack = factor(ifelse(PctBlack > 0.60, 1, 0)),
        HighRev= factor(ifelse(reviews >= 100, 1, 0)))


head(AirBnb.data)
glimpse(AirBnb.data)

```

# Research Question 1: What variables contribute to a customer's complete satisfaction with a hotel? 
### EDA
As seen in the summary statistics, the mean bedrooms, number for people the unit accommodates, price, Walk score, Transit score, and Bike Score for being completely satisfied (satisfaction = 1) and not completely satsified (satisfaction = 0) are all very similar. The only one that looks significantly different is the price of the unit. After looking at the boxplots, the means for bedrooms, transit score, and bike score appear to be very similar and the means for number of people the unit accommodates, price, and walk score look different. Therefore, I will fit a logistic model to each of these variables to see if they are significant in predicting the odds of whether of not a customer is completely satisfied. 

```{r}
AirBnb.data %>%
  group_by(satisfaction) %>%
  summarise(mean(bedrooms), 
            mean(accommodates), 
            mean(price), 
            mean(minstay), 
            mean(WalkScore), 
            mean(TransitScore), 
            mean(BikeScore))

ggplot(AirBnb.data, aes(x = satisfaction, y = bedrooms)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = accommodates)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = price)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = minstay)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = WalkScore)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = TransitScore)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = BikeScore)) + geom_boxplot(color = "black") + coord_flip()


```

### Fitting the Models for Research Question 1
* First I tested each variable (bedrooms, accommodates, price, minstay, WalkScore, TransitScore and BikeScore) to see if they were significant in predicting the odds that a customer is satisfied with their stay at their hotel. After fitting each model, it appeared that the significant variables were accommodates, price and WalkScore. Then, I decided to create a full model which would contain the additive model of the significant variables as well as their interactions. After the backwards elimination process, the full final model turned out to be: 

\[log(pi/(1−pi)) = \beta_0 + \beta_1(accommodates) + \beta_2(price) + \beta_3(WalkScore) + \beta_4(accommodates*WalkScore) + \beta_5(price*WalkScore)\]

```{r}

# Bedtooms
bedModel <- glm(satisfaction ~ bedrooms, family = "binomial", data = AirBnb.data)
summary(bedModel)

# Accomodations
accModel <- glm(satisfaction ~ accommodates, family = "binomial", data = AirBnb.data)
summary(accModel)

# Price
priceModel <- glm(satisfaction ~ price, family = "binomial", data = AirBnb.data)
summary(priceModel)

# Minimum stay
minstayModel <- glm(satisfaction ~ minstay, family = "binomial", data = AirBnb.data)
summary(minstayModel)

# Walk Score
minstayModel <- glm(satisfaction ~ WalkScore, family = "binomial", data = AirBnb.data)
summary(minstayModel)

# Transit Score
minstayModel <- glm(satisfaction ~ TransitScore, family = "binomial", data = AirBnb.data)
summary(minstayModel)

# Bike Score
minstayModel <- glm(satisfaction ~ BikeScore, family = "binomial", data = AirBnb.data)
summary(minstayModel)

```

```{r}
#Full Model 
fullModel <- glm(satisfaction ~ (accommodates + price + WalkScore)^2, family = "binomial", data = AirBnb.data)

fullFinalModel <- step(fullModel, direction = "backward")
summary(fullFinalModel)
```

* Given that this is a somewhat of a complicated model with six parameter, two of which include an interaction between two variables, I wanted to test if the additive model would be sufficient in predicting the odds that the customer would be satisfied with their stay. 

* Therefore: 

\[H_0: log(pi/(1−pi)) = \beta_0 + \beta_1(accommodates) + \beta_2(price) + \beta_3(WalkScore)\]
\[H_a: log(pi/(1−pi)) = \beta_0 + \beta_1(accommodates) + \beta_2(price) + \beta_3(WalkScore) + \beta_4(accommodates*WalkScore) + \beta_5(price*WalkScore)\]

* The results of the likelihood ratio test were as follows: 
The LRT test statistic is 10.847 and the p-value is 0.004411. Because the p-value of 0.004411 is less than $\alpha$ = 0.05, we can reject the null and say that the interactions terms are significant in this model and should not be removed. 
```{r}
redFinalModel <- glm(satisfaction ~ accommodates + price + WalkScore, family = "binomial", data = AirBnb.data)

lrtest(fullFinalModel, redFinalModel)
```

* Therefore, the final model remains: 

\[log(pi/(1−pi)) = \beta_0 + \beta_1(accommodates) + \beta_2(price) + \beta_3(WalkScore) + \beta_4(accommodates*WalkScore) + \beta_5(price*WalkScore)\]

* Interpretations: (log odds --> odds($e$ raised to each parameter))

* $\beta_0$ = -1.7094179 --> 0.1809711; the estimated odds of being completely satisfied (satisfaction = 1) is 0.1809711.  
* $\beta_1$ = -1.0877718 --> 0.3369664; the estimated decrease in odds of the customer being completely satisfied with their stay for every one-unit increase in accommodates is 0.3369664, holding all other variables at a fixed value. 
* $\beta_2$ = 0.0592719 --> 1.0610637; the estimated increase in odds of the customer being completely satisfied with their stay for every one-unit increase in price is 1.0610637, holding all other variables at a fixed value.
* $\beta_3$ = 0.0268196 --> 1.0285972; the estimated increase in odds of the customer being completely satisfied with their stay for every one-unit increase in WalkScore is 1.0285972, holding all other variables at a fixed value.
* $\beta_4$ = 0.0099619 --> 1.0100116; For every one-unit increase in acommodations, the effect of an additional unit of WalkScore is estimated to increase the odds of a customer being completely satisfied by 1.0100116. 
* $\beta_5$ = -0.0006085 --> 0.9993916; For every one-unit increase in price, the effect of an additional unit of WalkSCore is estimated to increase the odds of a customer being completely satisfied by 1.0100116. 

### Conclusion: 
* After looking at summary statistics and barplots, the seven variables I was interested in testing were bedrooms, accommodates, price, minstay, WalkScore, TransitScore, and BikeScore. I fit each variable in its own logistic regression model to see if they were significant in predicting the odds of complete satisfaction for customers. The final model created after backwards elimination contained the variables, accommodates, price, WalkScore, as well as the interactions accommodates * walkScore and price * WalkScore. To test if the model could be reduced to the just the additive model, I performed an LRT test; the results showed that the interactions were significant and should be kept in the model. 

# Research Question 2: When renting an Air BnB, to what extent to the racial composition of the neighborhood, quality of neighborhood for walking, number of people the unit is able to accomodate and number of reviews,  affect the price of the unit you choose?
### EDA
* Level 1 (price) predictors: overall_satisfaction, satisfaction, reviews, room_type, accommodates, bedrooms, minstay

* Level 2 (neighborhood) predictors: district, WalkScore, TransitScore, BikeScore, PctBlack, and HighBlack

* The summary statistics given show us how the means of WalkScore and accommodates,and the counts for HighBlack and HighRev all vary from neighborhood to neighborhood. When testing these variables in the model, we will be able to see how they affect the mean price in each neighborhood. 

* Furthermore, in all neighborhoods, it seems to be the pattern that if there is a high percentage of blacks in the neighborhood, the price of the unit will be lower than if there were a lower percentage of blacks. 

* There could be a possible neighborhood by HighRev interaction. Some neighborhoods such as West Lawn and Pullman have higher prices for units with a high number of reviews. Other neighborhoods, however, such as Logan Square and Jefferson Park have higher prices for units with less reviews. 
```{r fig.height=10, fig.width=7}

AirBnb.data %>%
  group_by(neighborhood) %>%
  summarise(mean(price),
            mean(WalkScore),
            mean(accommodates),
            count(HighBlack),
            count(HighRev))
          
```

```{r}
ggplot(data=AirBnb.data, aes(x=HighBlack,y=price)) + 
  geom_boxplot() + coord_flip() + ggtitle("HighBlack vs Price (all neighborhoods combined0")
```
```{r fig.height=10, fig.width=7}
ggplot(data=AirBnb.data, aes(x= HighBlack,y=price)) +  geom_dotplot(binaxis="y") +    coord_flip() + facet_wrap(~neighborhood,ncol=5) + ggtitle("HighBlack vs. Price")
```

```{r}
ggplot(data=AirBnb.data, aes(x=HighRev,y=price)) + 
  geom_boxplot() + coord_flip() + ggtitle("HighRev vs. Price (all neighborhoods combined)")
```

```{r fig.height=10, fig.width=7}
ggplot(data=AirBnb.data, aes(x= HighRev,y=price)) +  geom_dotplot(binaxis="y") +    coord_flip() + facet_wrap(~neighborhood,ncol=5) + ggtitle("HighBlack vs. Price")
```

### Fitting the Models for Research Question 2
##### Unconditional Means Model to asses amount of variation at each level
* $\alpha_0$ = 97.289
* $\sigma^2_u$ = 1099
* $\sigma^2$ = 6776
* ICC = (1099)/ (1009 + 6776) = 0.1411689
-> 14.12% ot total variability in price of units are attributed to difference among neighborhoods.

```{r}
modelA <- lmer(price ~ 1 + (1|neighborhood), data = AirBnb.data)
summary(modelA)
```


##### Random Sloped and Intercepts Model for level one covariates - HighRev and accommodates
* Level 1: 

\[Y_{ij}=a_i+b_i(HighRev) + c_i(accommodates) + \varepsilon_{ij}\]

* Level 2: 

\begin{align*}
a_i=\alpha_0 + u_i\\
b_i=\beta_0 + v_i\\
c_i=\gamma_0 + w_i
\end{align*}

* Composite Model: 

\[Y_{ij} = \alpha_0 + \beta_0(HighRev) + \gamma_0(accommodates) + u_i + v_i(HighREv) + w_i(accommodates) + \varepsilon_{ij}\]

where $\varepsilon_{ij}\sim N(0,\sigma^2)$, $u_i \sim N(0,\sigma^2_u)$, $v_i \sim N(0,\sigma^2_u)$, $w_i \sim N(0,\sigma^2_u)$. 
```{r}
modelB <- lmer(price ~ HighRev + accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
summary(modelB)

```

* $\hat{\alpha}_{0}$ = 29.603; The mean price of an Air Bnb is 29.603.
* $\hat{\beta}_{0}$ = -8.600; The mean decrease in price for units with high reviews is -8.600. 
* $\hat\gamma_{0}$ = 19.118; The mean increase in price for every one-unit increase in accommodation is 19.138. 
* $\hat{\sigma}_{2u}$= 41.06;  The variance in between-neighborhood deviations in price for units having a high number of reviews is 41.06.
* $\hat{\sigma}_{2v}$= 114.54;  The variance in between-person deviations in price for every increase in accommodations if is 114.54.
* $\hat{\sigma}_2$ = 3641.82, the estimated variance in within neighborhood deviations. 
* $\rho_{uv}$ = 0.-11; the correlation in an Air Bnb's price that has a high number of reviews and their differences in price for ever one-unit increase in accommodations.

##### Adding Level Two Covariates: HighBlack and WalkScore
* Level 1: 

\[Y_{ij}=a_i+b_i(HighRev) + c_i(accommodates) + \varepsilon_{ij}\]

* Level 2: 

\begin{align*}
a_i=\alpha_0 + \alpha_1(HighBlack) + \alpha_2(WalkScore) + u_i\\
b_i=\beta_0 + \beta_1(HighBlack) + \beta_2(WalkScore) + v_i\\
c_i=\gamma_0 + \gamma_1(HighBlack) +\gamma_2(WalkScore) + w_i
\end{align*}

* Composite Model: 

\[Y_{ij} = \alpha_0 + \alpha_1(HighBlack) + \alpha_2(WalkScore) + \beta_0(HighRev) + \beta_1(HighBlack*HighRev) + \beta_2(WalkScore* HighRev) + \gamma_0(accommodates) + \gamma_1(HighBlack*accommodates) + \gamma_2(WalkScore*accommodates) u_i + v_i(HighRev) + w_i(accommodates) + \varepsilon_{ij}\]

where $\varepsilon_{ij}\sim N(0,\sigma^2)$, $u_i \sim N(0,\sigma^2_u)$, $v_i \sim N(0,\sigma^2_u)$, $w_i \sim N(0,\sigma^2_u)$. 

```{r}
modelC <- lmer(price ~ HighBlack*HighRev + HighBlack*accommodates + WalkScore*HighRev + WalkScore*accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
summary(modelC)
``` 


* As seen above, none of the variables are significant in predicting price. Therefore, through backwards elimination, I would like to see if we can simplify the model further. After doing this, the final model is: 

\[Y_{ij} = \alpha_0 + \alpha_1(WalkScore) + \beta_0(accommodates) + \beta_1(Walkscore*accommodates)\]


```{r message=FALSE, warning=FALSE}
# Remove HighBlack*HighRev interaction
modelC_1 <- lmer(price ~ HighBlack + HighRev + accommodates + WalkScore + HighBlack*accommodates + WalkScore*HighRev + WalkScore*accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
# summary(modelC_1)

# Remove HighRev*WalkScore
modelC_2 <- lmer(price ~ HighBlack + HighRev + accommodates + WalkScore + HighBlack*accommodates + WalkScore*accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
#summary(modelC_2)

# Remove HighBlack*accommodates
modelC_3 <- lmer(price ~ HighBlack + HighRev + accommodates + WalkScore + WalkScore*accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
#summary(modelC_3)

# Remove HighBlack
modelC_4 <- lmer(price ~ HighRev + accommodates + WalkScore + WalkScore*accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
#summary(modelC_4)

# Remove HighRev
modelC_4 <- lmer(price ~ accommodates + WalkScore + WalkScore*accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
summary(modelC_4)

```

* However, would the additive model, with all four variables, better explain price variation within neighborhoods? First, I will compare the full model, model C, to the additive model, model D. Then I will compare the additive model to the model given from backwards elimination. 
* Comparing model C and D: 
The LRT statistic is 11.57 and the p-value is 0.02085. Because the p-value of 0.02085 is less than $\alpha$ = 0.05, we reject the null hypothesis and have sufficient evidence to say that modelC, the full model, is a better model. 
* Comparing model D and model C_4
The LRT statistic is 0 and the p-value is 1. Because the p-value of 1 is greater than $\alpha$ = 0.05, we fail to reject the null hypothesis that the further reduced model, modelC_4, is a sufficient model. *These test opose themselves, the first test tells us that the full model is better than the additive model, but the second test tells us that the further reduced model is better than the additive model. Therefore, it would be good to check the AIC and BIC for each model in order to make a decision on the final model.
```{r}
modelD <- lmer(price ~ HighBlack + WalkScore + HighRev + accommodates + (HighRev + accommodates|neighborhood), data = AirBnb.data)
summary(modelD)

anova(modelC, modelD)

anova(modelD, modelC_4)
```
*The AIC shows that modelC is the better model, but the BIC shows that modelC_4 is better. Therefore, I choose to stand by modelC because I would rather take the chance of some predictors not explaining price versus missing predictors that could be potentially significant. 
```{r}
#Full model
AIC(modelC)
BIC(modelC)

#Additive model
AIC(modelD)
BIC(modelD)

#Further reduced model
AIC(modelC_4)
BIC(modelC_4)
```

* Therefore the final model and its interpretations are:  

\[Y_{ij} = \alpha_0 + \alpha_1(HighBlack) + \alpha_2(WalkScore) + \beta_0(HighRev) + \beta_1(HighBlack*HighRev) + \beta_2(WalkScore* HighRev) + \gamma_0(accommodates) + \gamma_1(HighBlack*accommodates) + \gamma_2(WalkScore*accommodates) u_i + v_i(HighRev) + w_i(accommodates) + \varepsilon_{ij}\]

* $\hat{\alpha}_{0}$ = 3.1165; The estimated mean price of an Air Bnb when there is a low percentage of black population, WalkScore is 0, a low number of reviews and accommodates zero people is 3.1165.

* $\hat{\alpha}_{1}$ = 21.3451; The estimated mean price of an Air Bnb when there is a high black population, controlling for the other variables, is 21.3451 higher than neighborhoods with low black populations. 

* $\hat{\alpha}_{2}$ = 0.3025; The estimated mean increase in price of an Air Bnb for every one-unit increase in WalkScore, controlling for the other varibalesis, is 0.3025.

* $\hat{\beta}_{0}$ = 70.4237; The estimated mean difference in price for units with high reviews is 70.4237 more than units with a low number of reviews, controlling for the other variables. 

* $\hat{\beta}_{1}$ = -1.4321; The estimated mean difference in price for Air BnB's with a high black population percentage is 1.4321 dollars lower than Air BnB's with low black population percentage, when the unit has high reviews and the other variables are controlled.

* $\hat{\beta}_{2}$ = -0.9429; The estimated mean difference in price for Air Bnb's with a high number of reviews is 0.9429 less than those with low reviews, for every one-unit increase in Walk Score and controlling for other variables. 

* $\hat{\gamma}_{0}$ = -21.3982; The estimated mean decrease in price for Air Bnb's for every one-unit increase in accomodates is 21.3982. 

* $\hat{\gamma}_1$ = -9.2030; The estimated mean difference in price for Air Bnb's with a high black population percentage is 9.2030 less than Air Bnb's with a low black population percentage, for every one-unit incresae in accommodates and controlling for other variables. 

* $\hat{\gamma}_2$ = 0.5224; When WalkScore increases by one-unit, the effect of a one-unit increase in accommodates is estimated to increase the mean price of an Air Bnb by 0.5224. 

* $\hat{\sigma}_{u}$= 17.11;  The variance in between-neighborhood deviations in price for Air Bnb's with a high number of reviews is 17.11.

* $\hat{\sigma}_{v}$= 83.27;  The variance in between-neighborhood deviations in price for an Air Bnb for every one-unit increase in accommodates is 83.27. 

* $\rho_{uv}$ = -0.81; the correlation in an Air Bnb's price that has a high number of reviews and their differences in price for ever one-unit increase in accommodations.

### Conclusion: 
* Given that my dad is always worried about the price of a hotel, I wanted to look into what factors contribute to the price of an Air Bnb. The four variables that I chose to look at were HighBlack (greater than 60 % of the population in neighborhood is black), HighRev (Air Bnb has more than 100 reviews posted), accommodates (how many the Air Bnb can accommodate), and WalkScore (how "walkable" the neighborhood is). After fitting the full model, I wanted to see if the model could be reduced further. However, after comparing the full model to the additive model, the full model turned out to be better. However, when comparing the additive model to the even further-reduced model (resulting from backwards elimination), the further-reduced model turned out to be better. Therefore, after looking at the AIC and BIC of each model, I decided to keep the full model in hopes of not losing any variables that could potentially explain the estimated mean price. 

