---
title: "Air BnB Data Analysis"
author: Shravani Samala
date: "4/2/2020 "
output: html_document
---

```{r, message=F, warning=F}
# front-matter
rm(list = ls()) #clear the workspace

library(ggplot2)
library(ggformula)
library(corrplot)
library(dplyr)
library(lmerTest)
library(lme4)
library(knitr)
library(tidyverse)
library(readr)
library(mosaic)
library(data.table)
library(VGAM)
library(lmtest)

```
# Introduction

*overall_satisfaction = rating on a 0-5 scale.
*satisfaction = 1 if overall_satisfaction is 5, 0 otherwise
*price = price for one night (in dollars)
*reviews = number of reviews posted
*room_type = Entire home/apt, Private room, or Shared room
*accommodates = number of people the unit can hold
*bedrooms = number of bedrooms
*minstay = minimum length of stay (in days)
*neighborhood = neighborhood where unit is located (1 of 43)
*district = district where unit is located (1 of 9)
*WalkScore = quality of the neighborhood for walking (0-100)
*TransitScore = quality of the neighborhood for public transit (0-100)
*BikeScore = quality of the neighborhood for biking (0-100)
*PctBlack = proportion of black residents in a neighborhood
*HighBlack = 1 if PctBlack above .60, 0 otherwise
```{r}
myFile2 <- "https://raw.githubusercontent.com/proback/BYSH/master/data/airbnb.csv"
AirBnb.data <- read_csv(myFile2)

head(AirBnb.data)
```


```{r}
AirBnb.data<-
AirBnb.data%>%
  rename(hotelID = X1)%>%
  mutate(priceHigh = factor(ifelse(price >= 100,1,0)),
         satisfaction = factor(ifelse(overall_satisfaction == 5.0, 1, 0))) # 1 means high price, 0 means low price
head(AirBnb.data)
glimpse(AirBnb.data)
View(AirBnb.data)
```

# Research Question 1: What variables contribute to a customer's overall satisfaction with a hotel? 
```{r}

ggplot(AirBnb.data, aes(x = satisfaction, y = bedrooms)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = accommodates)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = price)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = minstay)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = WalkScore)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = TransitScore)) + geom_boxplot(color = "black") + coord_flip()
ggplot(AirBnb.data, aes(x = satisfaction, y = BikeScore)) + geom_boxplot(color = "black") + coord_flip()






```
```

Are the number of bedrooms and the number of people a hotel can accomodate relate to the odds of a hotel having a high price?
```{r}
model.PriceHigh <- glm(priceHigh ~ bedrooms + accomodates + , family = binomial, data = data.cereal)
summary(mod.HighSugar)
```

```{r}
modelA <- lmer(price ~ 1 + (1|neighborhood), data = AirBnb.data)
summary(modelA)

```