---
title: "Abortion Analysis"
author: Shravani Samala
date: "4/2/2020 "
output: html_document
---

# Introduction
```{r, message=F, warning=F}
# front-matter
rm(list = ls()) #clear the workspace

library(ggplot2)
library(dplyr)
library(lmerTest)
library(lme4)
library(knitr)
library(tidyverse)
library(readr)
library(mosaic)
library(data.table)

```

Introduction: 
Five regions: West, Midwest, Southwest, Southeast, and Northeast. Randomly choose four states from each region and five counties from each state. 
West: Oregon, Utah, Montana, and Colorado
Midwest: Ohio, South Dakota, and Michigan
Southwest: Arizona, New Mexico, Texas, and Oklahoma
Southeast: Arkansas, North Carolina, Alabama, and Virginia
Northeast: Pennsylvania, Massachusetts, Vermont, and Connecticut


```{r}
# Import data from github
myFile <- "https://raw.githubusercontent.com/MultiLevelAnalysis/Datasets-third-edition-Multilevel-book/master/chapter%207/Safety/safety.dat"

data <- fread(myFile)

#abortion <- read_csv(myFile)

head(data)
glimpse(data)
View(data)

StreetData <-
  data %>%
  rename(street = V1,
         personID = V2, 
         safety = V7)
  
StreetData

```

```{r}
# Clean up the data


```

